Replacing standard galaxy profiles with \\ mixtures of Gaussians

### 2012-10-23 documents/mixtures/mixture_models.tex
notes to self

% This file is part of the Tractor project.
% Copyright 2012 David W. Hogg (NYU) and Dustin Lang (Princeton->CMU).
% All rights reserved.

% to-do
% -----
% - HOGG: can we drop the regularization in the badness and still succeeed?  If so, let's do that (and drop it in the text).
% - HOGG: need Bundy reference
% - HOGG: make what I say about M=6 lux and M=8 lux true -- ie, edit the Tractor code!
% - HOGG: send to Lupton and Bundy for comments
% - DSTN: can we be more specific about how Bosch is using them? I would like to cite users.
% - HOGG: Cite paper mentioned by Bosch?
% - HOGG: should we make Machine Readable Tables for PASP?

@@

Applying this PSF to an exp or dev galaxy is slightly more
complicated, because the galaxy has not just a flux $S_g$ and a
central position $\pos_g$; it also has a shape.  Because we are only
considering these simple galaxies, we are only permitting ellipsoidal
shapes, which can be represented by a semi-major axis $a$, a
semi-minor axis $b$, and a position angle $\phi$, or equivalently by
eigenvalues $a, b$ and eigenvectors $\uv_1, \uv_2$, or equivalently by
an affine transformation $\affine_g$ that takes a circle to the
relevant ellipse (and is a therefore a general representation of an
ellipse; it is also the matrix square root of the symmetric variance
tensor describing the ellipse).  The galaxy is distorted by this
affine transformation \emph{prior} to PSF convolution, so the
focal-plane image (PSF-convolved intensity field) for a general (say)
$\exp$ galaxy is given by
\begin{eqnarray}\displaystyle
I(\pos\given\exp,S_g,\pos_g,\affine_g) &=& \sum_{k=1}^K \sum_{m=1}^{M^{\exp}} S_g\,a^{\exp}_m\,p_k\,\normal(\pos\given\pos_g+\mean_k,\var_{gm}+\var_k)
\\
\var_{gm} &\equiv& \affine_g\cdot\var^{\exp}_m\cdot\transpose{\affine_g}
\\
\affine_g &=& \left[a\,\uv_1 , b\,\uv_2 \right]
\quad ,
\end{eqnarray}
where $a$ and $b$ are the major and minor axis lengths of the galaxy
ellipse (in appropriate units) and $\uv_1$ and $\uv_2$ are the
eigenvectors on the sky pointing in the major-axis and minor-axis
directions respectively.  Implicitly all vectors are two-dimensional
column vectors, and $\affine_g$ is a $2\times 2$ affine transformation
matrix that contains the ``shape'' (position angle, major-axis, and
ellipticity) information about the galaxy.  The $\dev$, $\luv$, and
$\lux$ cases are all similar.  Note the important and key result of
this \documentname, to wit, that a MoG galaxy model (with $M$
components) convolved with a MoG PSF model (with $K$ components)
yields a MoG model image (with $[M\,K]$ components).

@@

In \figurename~\ref{fig:example} we show how we are using these MoG
approximations in \project{The Tractor}---a generative modeling
framework for measuring astronomical objects---to render PSF-convolved
galaxy images.

@@

\begin{thebibliography}{70}
\bibitem[Bendinelli(1991)]{bendinelli}
Bendinelli,~O., 1991, \apj, 366, 599
\bibitem[Bendinelli \& Parmeggiani(1995)]{bendinelli2}
Bendinelli,~O. \& Parmeggiani,~G., 1995, \aj, 109, 572
\bibitem[Bosch(2010)]{bosch}
Bosch,~J., 2010, \aj, 140, 870
\bibitem[Bovy \etal(2011a)]{xd}
Bovy,~J., Hogg,~D.~W., \& Roweis,~S., 2011, Ann. Appl. Stat., 5, 1657
\bibitem[Bovy \etal(2011b)]{xdqso}
Bovy,~J., \etal, 2011, \apj, 729, 141
\bibitem[Bovy \etal(2012)]{xdqsoz}
Bovy,~J. \etal, 2012, \aj, 749, 41
\bibitem[Bundy \etal(2012)]{bundy}
Bundy,~K.~??
\bibitem[Cappellari(2002)]{cappellari}
Cappellari,~M., 2002, \mnras, 333, 400
\bibitem[Ciotti \& Bertin(1999)]{ciotti}
Ciotti,~L. \& Bertin,~G., 1999, \aa, 352, 447
\bibitem[de~Vaucouleurs(1948)]{dev}
de~Vaucouleurs,~G., 1948, Annales d'Astrophysique, 11, 247
\bibitem[Emsellem \etal(1994)]{emsellem}
Emsellem,~E., Monnet,~G., Bacon,~R., \& Nieto,~J.-L., 1994, \aap, 285, 739
\bibitem[Lupton \etal(2001)]{lupton}
Lupton,~R., Gunn,~J.~E., Ivezic,~Z., Knapp,~G.~R., Kent,~S.~M., \& Yasuda,~N., 2001, ASPC, 238, 269
\bibitem[Peng \etal(2002)]{galfit}
Peng,~C.~Y., Ho, L.~C., Impey,~C.~D., \& Rix,~H.-W., 2002, \aj, 124, 266
\bibitem[S{\'e}rsic(1963)]{ser}
S{\'e}rsic,~J.~L., 1963, Boletin de la Asociacion Argentina de Astronomia La Plata Argentina, 6, 41
\end{thebibliography}

>>>>

% This file is part of The Tractor.
% Copyright 2012 David W. Hogg (NYU) and Dustin Lang (Princeton, CMU).
% All rights reserved.

% to-do
% -----
% - HOGG: need Bundy reference
% - HOGG: send to Rix and Peng and Lupton and Bundy for comments
% - HOGG: submit!
% - HOGG: can we drop the regularization in the badness and still succeeed?  If so, let's do that (and drop it in the text).
% - HOGG: make what I say about M=6 lux and M=8 lux true -- ie, edit the Tractor code!
% - DSTN: can we be more specific about how Bosch is using them? I would like to cite users.
% - HOGG: should we make Machine Readable Tables for PASP?  Yes.

@@

Applying this PSF to an exp or dev galaxy is slightly more
complicated, because the galaxy has not just a flux $S_g$ and a
central position $\pos_g$; it also has a shape.  Because we are only
considering these simple galaxies, we are only permitting ellipsoidal
shapes, which can be represented by a semi-major axis $a$, a
semi-minor axis $b$, and a position angle $\phi$, or equivalently by
eigenvalues $a, b$ and eigenvectors $\uv_1, \uv_2$, or equivalently by
an affine transformation $\affine_g$ that takes a circle to the
relevant ellipse (and is a therefore a general representation of an
ellipse; it is also the matrix square root of the symmetric variance
tensor describing the ellipse).  The galaxy is distorted by this
affine transformation \emph{prior} to PSF convolution, so the
focal-plane image (PSF-convolved intensity field) for a general (say)
$\exp$ galaxy is given by
\begin{eqnarray}\displaystyle
I(\pos\given\exp,S_g,\pos_g,\affine_g) &=& \sum_{k=1}^K \sum_{m=1}^{M^{\exp}} S_g\,a^{\exp}_m\,p_k\,\normal(\pos\given\pos_g+\mean_k,\var_{gm}+\var_k)
\\
\var_{gm} &\equiv& \affine_g\cdot\var^{\exp}_m\cdot\transpose{\affine_g}
\\
\affine_g &=& \left[a\,\uv_1 , b\,\uv_2 \right]
\quad ,
\end{eqnarray}
where $a$ and $b$ are the major and minor axis lengths of the galaxy
ellipse (in appropriate units) and $\uv_1$ and $\uv_2$ are the
eigenvectors on the sky pointing in the major-axis and minor-axis
directions respectively.  Implicitly all vectors are two-dimensional
column vectors, and $\affine_g$ is a $2\times 2$ affine transformation
matrix that contains the ``shape'' (position angle, major-axis, and
ellipticity) information about the galaxy.  The $\dev$, $\luv$, and
$\lux$ cases are all similar.  Note the important and key result of
this \documentname, to wit, that a MoG galaxy model (with $M$
components) convolved with a MoG PSF model (with $K$ components)
yields a MoG model image (with $[M\,K]$ components).  In
\figurename~\ref{fig:example} we show how we are using these MoG
approximations in \project{The Tractor}---a generative modeling
framework for measuring astronomical objects---to render PSF-convolved
galaxy images.

@@

@@

\begin{thebibliography}{70}
\bibitem[Bendinelli(1991)]{bendinelli}
Bendinelli,~O., 1991, \apj, 366, 599
\bibitem[Bendinelli \& Parmeggiani(1995)]{bendinelli2}
Bendinelli,~O. \& Parmeggiani,~G., 1995, \aj, 109, 572
\bibitem[Bosch(2010)]{bosch}
Bosch,~J., 2010, \aj, 140, 870
\bibitem[Bovy \etal(2011a)]{xd}
Bovy,~J., Hogg,~D.~W., \& Roweis,~S., 2011, Ann. Appl. Stat., 5, 1657
\bibitem[Bovy \etal(2011b)]{xdqso}
Bovy,~J., \etal, 2011, \apj, 729, 141
\bibitem[Bovy \etal(2012)]{xdqsoz}
Bovy,~J. \etal, 2012, \aj, 749, 41
\bibitem[Bundy \etal(2012)]{bundy}
Bundy,~K. \etal, \aj, in preparation (arXiv:HOGG)
\bibitem[Cappellari(2002)]{cappellari}
Cappellari,~M., 2002, \mnras, 333, 400
\bibitem[Ciotti \& Bertin(1999)]{ciotti}
Ciotti,~L. \& Bertin,~G., 1999, \aa, 352, 447
\bibitem[de~Vaucouleurs(1948)]{dev}
de~Vaucouleurs,~G., 1948, Annales d'Astrophysique, 11, 247
\bibitem[Emsellem \etal(1994)]{emsellem}
Emsellem,~E., Monnet,~G., Bacon,~R., \& Nieto,~J.-L., 1994, \aap, 285, 739
\bibitem[Lupton \etal(2001)]{lupton}
Lupton,~R., Gunn,~J.~E., Ivezic,~Z., Knapp,~G.~R., Kent,~S.~M., \& Yasuda,~N., 2001, ASPC, 238, 269
\bibitem[Peng \etal(2002)]{galfit}
Peng,~C.~Y., Ho, L.~C., Impey,~C.~D., \& Rix,~H.-W., 2002, \aj, 124, 266
\bibitem[S{\'e}rsic(1963)]{ser}
S{\'e}rsic,~J.~L., 1963, Boletin de la Asociacion Argentina de Astronomia La Plata Argentina, 6, 41
\end{thebibliography}
